Reproducible Workflow on a Public Cloud for Computational Fluid Dynamics

### 2019-04-10 tex/MesnardoBarba2019.tex
final edits results section

Our research lab is interested in understanding the aerodynamics of flying animals via CFD simulations.
One of our applications deals with the aerodynamics of a snake species, \textit{Chrysopelea Paradisi}, that lives in South-East Asia.
The arboreal reptile has the remarkable capacity to turn its entire body into a wing and glide over several meters\cite{socha_2011}.
The so-called ``flying snake'' jumps from tree branches, undulates in the air, and is able to produce lift by expanding its ribcage to flatten its ventral surface (morphing its normally circular cross-section into a triangular shape).

To study the flow around the flying snake, we develop a in-house CFD software called PetIBM\cite{chuang_et_al_2018}, a open-source toolbox that solves the two- and three-dimensional incompressible Navier-Stokes equations using a projection method (seen as an approximate block-LU decomposition of the fully discretized equations\cite{perot_1993}) and an Immersed-Boundary Method (IBM).
Within this framework, the fluid equations are solved over a extended grid that does not conform to the surface of the immersed body.
To model the presence of the body, the momentum equation is augmented with a forcing term that is activated in the vicinity of the immersed boundary.
This technique allows the use of simple fixed structured Cartesian grids to solve the equations.
PetIBM also implements Immersed Boundary Methods (IBMs) that fit into the projection method and, in the present study, we use an implementation of the IBM proposed in \cite{li_et_al_2016}.
PetIBM runs on distributed-memory architectures using the efficient data structures and parallel routines from the PETSc library.
The software also implements the possibility to solve linear systems on multiple GPU devices distributed across the nodes with the NVIDIA AmgX and AmgXWrapper\cite{chuang_barba_2017}.
One of the requirements to produce reproducible computational results is the make the code available.
In that regard, PetIBM is open source, version-controlled on GitHub,\footnote{PetIBM (version 0.4): \url{https://github.com/barbagroup/PetIBM}} and developed under the permissive BSD-3 clause license.

@@

With Batch Shipyard, we submitted a job on Azure Batch to compute the two-dimensional flow around an anatomically accurate sectional shape of the gliding snake.
The cross-section has a chord-length of $1$ and forms a $35$-degree angle of attack with the incoming freestream flow.
The Reynolds number, based on the freestream speed, the body chord-length, the kinematic viscosity, is set to $Re=2000$.
The immersed boundary is centered in a $30c \times 30c$ computational domain that contains just above $2.9$ million cells.
The grid is kept uniform with the highest resolution in the vicinity of the body and stretched to the external boundaries with a constant ratio (see Table \ref{tab:grid_specs} for details about the grid).
(The immersed boundary is discretized with the same resolution than the background fluid grid.)
A convective condition was used at the outlet boundary while freestream conditions were enforced on the three other boundaries.
The job was submitted to a pool of two \texttt{NC24r} nodes, using $12$ MPI processes and $2$ GPU devices per node.
The task completed $200,000$ time steps (i.e., $80$ time units of flow simulation with time-step size $\Delta t = 0.0004 c / U_\infty$) in just above $7$ hours.

Fig. \ref{fig:force_coefficients} shows the history of the force coefficients on the two-dimensional cross-section.
The lift coefficient only maintains its maximum mean value during the early stage of the simulation (up to $40$ time units).
Between $40$ and $50$ time units, the mean value starts to drop.
The time-averaged force coefficients (between $40$ and $80$ time units) are reported in Table \ref{tab:force_coefficients}.
Fig. \ref{fig:wz_2d} shows snapshots of the vorticity field after $20$, $44$, $45$, and $80$ time unit of flow simulation.
After $20$ time units, the vortices shed from the snake section are almost aligned in the near wake (with a slight deflection towards the lower part of the domain).
Snapshots of the vorticity at time units $44$ and $45$ show that the initial alignment is perturbed by vortex-merging events (same-sign vortices merging together to form a stronger one).
Following that, the wake signature is altered for the rest of the simulation: vortices are not aligned anymore, the wake becomes wider (leading to lower aerodynamic forces) with a 1S+1P vortex signature (a single clockwise-rotating vortex on the upper part and a vortex pair on the lower part).

@@

Although the two-dimensional simulations gave us some insights into to flow dynamics happening in the behind the snake, we know that at this Reynolds number ($Re = 2000$), three-dimensional structures will develop in the wake.
Thus, we submitted jobs to Azure Batch to perform direct numerical simulation of the three-dimensional flow around snake model: a cylinder with the same anatomically accurate cross-section.
The computational domain is extended in the $z$-direction over a length of $3.2c$ and the grid now contains about $46$ million cells (with a uniform discretization in the $z$-direction; see Table \ref{tab:grid_specs}).
The grid is kept uniform in the $z$-direction and we enforce periodic conditions at the front and back boundaries.
The simulation ran in a Docker container on Azure Batch using Batch Shipyard.
The job was submitted to a pool of two \texttt{NC24r} nodes using $24$ MPI processes and $4$ GPU devices per node.
The task completed $100,000$ time steps ($100$ time units with a time-step size $\Delta t = 0.01 c / U_\infty$) in about $5$ days and $16$ hours.

Fig. \ref{fig:force_coefficients} compares the history of the force coefficients between the two- and three-dimensional configurations.
The force coefficients resulting from the two-dimensional simulation of the snake are higher than those obtained for the snake cylinder; a relative difference of $+37.9\%$ and $+14.4\%$ for the time-averaged drag and lift coefficient, respectively (see Table \ref{tab:force_coefficients}).
Two-dimensional computational simulations of fundamentally three-dimensional flows lead to incorrect estimation of the force coefficients\cite{mittal_balachandar_1995}.

Fig. \ref{fig:wz_avg_3d} shows the instantaneous spanwise vorticity averaged along the $z$-direction after $80$ and $100$ time units.
Compared to the snapshots from the two-dimensional simulation, we observe that (1) free-shear layers roll up into vortices further away from the snake body and (2) the von Karman street exhibits a narrower wake.
We also note the presence of an unsteady recirculation region just behind the snake cylinder and alternating regions of positive and negative cross-flow velocity showing the presence of von Karman vortices (Fig. \ref{fig:ux_uy_xz_plane_3d}).
Fig. \ref{fig:qcrit_wx_3d} shows a side-view of the isosurfaces of the Q-criterion after $100$ time units and highlights the complexity of the three-dimensional turbulent wake generated by the snake model.

\begin{figure}[!h]
    \centering
    \includegraphics[width=8cm]{figures/forceCoefficientsCompare2D.pdf}
    \caption{History of the force coefficients obtained with two- and three-dimensional simulations at Reynolds number $2000$ for a snake cross-section with a $35$-degree angle of attack.}
    \label{fig:force_coefficients}
\end{figure}

\begin{table*}[!h]
    \renewcommand{\arraystretch}{1.5}
    \caption{Details about the computational grids used for the snake simulations with PetIBM. Distances are expressed in terms of chord-length units.}
    \label{tab:grid_specs}
    \centering
        \begin{tabular}{clllll}
        Case & Domain & Uniform region & Smallest cell-width & Stretching ratio & Size \\
        \hline
        2D & $30 \times 30$ & $\left[ -0.52, 3.48 \right] \times \left[ -2, 2 \right]$ & $0.004 \times 0.004$ & $1.01$ & $1704 \times 1706$ \\
        3D & $30 \times 30 \times 3.2$ & $\left[ -0.52, 3.48 \right] \times \left[ -2, 2 \right] \times \left[ 0, 3.2 \right]$ & $0.008 \times 0.008 \times 0.08$ & $1.01$ & $1071 \times 1072 \times 40$ \\
        \hline
    \end{tabular}
\end{table*}

@@

\begin{figure}[!h]
    \centering
    \includegraphics[width=8cm]{figures/wz_avg_multi_contourf.png}
    \caption{Filled contour of the spanwise-averaged z-component of the vorticity ($-5 \leq w_z c / U_\infty \leq 5$) field after $80$ and $100$ time units of flow simulation with PetIBM for the snake cylinder with a cross-section at a $35$-degree angle of attack and Reynolds number $2000$.}
    \label{fig:wz_avg_3d}
\end{figure}

\begin{figure}[!h]
    \centering
    \includegraphics[width=8cm]{figures/ux_uy_xz_plane.png}
    \caption{Filled contour of the streamwise velocity (top) and cross-flow velocity (bottom) behind the snake cylinder in the $x/z$ plane at $y/c=-0.2$ in the wake of the snake cylinder with a $35$-degree angle of attack at Reynolds number $2000$ after $100$ time-units of flow simulations. There are 52 contours from $-1.0$ to $1.0$. The grey area shows a projection of the snake cylinder in the $x/z$ plane. The solid black line defines the contour with $u_x = 0$ (top) and $u_y = 0$ (bottom).}
    \label{fig:ux_uy_xz_plane_3d}
\end{figure}

>>>>

Our research lab is interested in understanding the aerodynamics of flying animals via CFD simulations.
One of our applications deals with the aerodynamics of a snake species, \textit{Chrysopelea Paradisi}, that lives in South-East Asia.
This arboreal reptile has the remarkable capacity to turn its entire body into a wing and glide over several meters\cite{socha_2011}.
The so-called ``flying snake'' jumps from tree branches, undulates in the air, and is able to produce lift by expanding its ribcage to flatten its ventral surface (morphing its normally circular cross-section into a triangular shape).

To study the flow around the flying snake, we developed a CFD software called PetIBM\cite{chuang_et_al_2018}, an open-source toolbox that solves the two- and three-dimensional incompressible Navier-Stokes equations using a projection method (seen as an approximate block-LU decomposition of the fully discretized equations\cite{perot_1993}) and an immersed-boundary method (IBM).
Within this framework, the fluid equations are solved over a extended grid that does not conform to the surface of the immersed body.
To model the presence of the body, the momentum equation is augmented with a forcing term that is activated in the vicinity of the immersed boundary.
This technique allows the use of simple fixed structured Cartesian grids to solve the equations.
PetIBM also implements immersed boundary methods (IBMs) that fit into the projection method; in the present study, we use the IBM scheme proposed in \cite{li_et_al_2016}.
PetIBM runs on distributed-memory architectures using the efficient data structures and parallel routines from the PETSc library.
The software also implements the possibility to solve linear systems on multiple GPU devices distributed across the nodes with the NVIDIA AmgX library and our AmgXWrapper\cite{chuang_barba_2017}.
One of the requirements for reproducible computational results is to make the code available under a public license (allowing reuse and modification by others).
In that regard, PetIBM is open source, version-controlled on GitHub,\footnote{PetIBM (version 0.4): \url{https://github.com/barbagroup/PetIBM}} and shared under the permissive (non copy-left) BSD-3 clause license.

@@

We submitted a job on Azure Batch (via Batch Shipyard,) to compute the two-dimensional flow around an anatomically accurate sectional shape of the gliding snake.
The cross-section has a chord-length $c=1$ and forms a $35$-degree angle of attack with the incoming freestream flow.
The Reynolds number, based on the freestream speed, the body chord-length, the kinematic viscosity, is set to $Re=2000$.
The immersed boundary is centered in a $30c \times 30c$ computational domain that contains just over $2.9$ million cells.
The grid is kept uniform with the highest resolution in the vicinity of the body and stretched to the external boundaries with a constant ratio (see Table \ref{tab:grid_specs} for details about the grid).
For the immersed boundary, the same grid resolution than the background fluid grid was applied.
A convective condition was used at the outlet boundary while freestream conditions were enforced on the three other boundaries.
The job was submitted to a pool of two \texttt{NC24r} nodes, using $12$ MPI processes and $2$ GPU devices per node.
It completed $200,000$ time steps (i.e., $80$ time units of flow simulation with time-step size $\Delta t = 0.0004 c / U_\infty$) in just above $7$ hours (wall-clock time).

Fig. \ref{fig:force_coefficients} shows the history of the force coefficients on the two-dimensional cross-section.
The lift coefficient only maintains its maximum mean value during the early stage of the simulation (up to $40$ time units).
Between $40$ and $50$ time units, the mean value starts to drop.
The time-averaged force coefficients (between $40$ and $80$ time units) are reported in Table \ref{tab:force_coefficients}.
Fig. \ref{fig:wz_2d} shows snapshots of the vorticity field after $20$, $44$, $45$, and $80$ time units of flow simulation.
After $20$ time units, the vortices shed from the snake section are almost aligned in the near wake (with a slight deflection towards the lower part of the domain).
Snapshots of the vorticity at time units $44$ and $45$ show that the initial alignment is perturbed by vortex-merging events (same-sign vortices merging together to form a stronger one).
Following that, the wake signature is altered for the rest of the simulation: vortices are not aligned anymore, the wake becomes wider (leading to lower aerodynamic forces) with a 1S+1P vortex signature (a single clockwise-rotating vortex on the upper part and a vortex pair on the lower part).

@@

Although the two-dimensional simulations can give us some insights into to flow dynamics happening in the wake behind the snake, we know that at this Reynolds number ($Re = 2000$), three-dimensional structures will develop in the wake.
We submitted jobs on Azure Batch to perform direct numerical simulation of the three-dimensional flow around a snake model: a cylinder with the an anatomically accurate cross-section.
The computational domain is extended in the $z$-direction over a length of $3.2c$ and the grid now contains about $46$ million cells (with a uniform discretization in the $z$-direction; see Table \ref{tab:grid_specs}).
The grid is kept uniform in the $z$-direction and we enforce periodic conditions at the front and back boundaries.
The simulation ran in a Docker container on Azure Batch using Batch Shipyard.
The job was submitted to a pool of two \texttt{NC24r} nodes using $24$ MPI processes and $4$ GPU devices per node.
The task completed $100,000$ time steps ($100$ time units with a time-step size $\Delta t = 0.01 c / U_\infty$) in about $5$ days and $16$ hours.

Fig. \ref{fig:force_coefficients} compares the history of the force coefficients between the two- and three-dimensional configurations.
The force coefficients resulting from the two-dimensional simulation of the snake are higher than those obtained for the snake cylinder; we computed a relative difference of $+37.9\%$ and $+14.4\%$ for the time-averaged drag and lift coefficient, respectively (see Table \ref{tab:force_coefficients}).
Two-dimensional computational simulations of fundamentally three-dimensional flows lead to incorrect estimation of the force coefficients, as is well known\cite{mittal_balachandar_1995}.

Fig. \ref{fig:wz_avg_3d} shows the instantaneous spanwise vorticity averaged along the $z$-direction after $80$ and $100$ time units.
Compared to the snapshots from the two-dimensional simulation, we observe that (1) free-shear layers roll up into vortices further away from the snake body and (2) the von Karman street exhibits a narrower wake than in the two-dimensional simulations.
We also note the presence of an unsteady recirculation region just behind the snake cylinder, and alternating regions of positive and negative cross-flow velocity showing the presence of von Karman vortices (Fig. \ref{fig:ux_uy_xz_plane_3d}).
Fig. \ref{fig:qcrit_wx_3d} shows a side-view of the isosurfaces of the Q-criterion after $100$ time units and highlights the complexity of the three-dimensional turbulent wake generated by the snake model.

The selection of results presented here corresponds to a typical set of experiments included in a CFD study. 
A comprehensive study might include several similar sets, leading to additional insights about the flow dynamics, and clues to future avenues of research.
We include this selection here to represent a typical workflow, and combine our discussion with a meaningful analysis of the costs associated with running CFD studies in a public cloud.

\begin{figure}
    \centering
    \includegraphics[width=8cm]{figures/forceCoefficientsCompare2D.pdf}
    \caption{History of the force coefficients obtained with two- and three-dimensional simulations at Reynolds number $2000$ for a snake cross-section with a $35$-degree angle of attack. As expected, the two-dimensional simulations over-estimate the force coefficients of what is essentially a three-dimensional flow problem.}
    \label{fig:force_coefficients}
\end{figure}

\begin{table*}
    \renewcommand{\arraystretch}{1.5}
    \caption{Details about the computational grids used for the snake simulations with PetIBM. Distances are expressed in terms of chord-length units.}
    \label{tab:grid_specs}
    \centering
        \begin{tabular}{clllll}
        Case & Domain & Uniform region & Smallest cell-width & Stretching ratio & Size \\
        \hline
        2D & $30 \times 30$ & $\left[ -0.52, 3.48 \right] \times \left[ -2, 2 \right]$ & $0.004 \times 0.004$ & $1.01$ & $1704 \times 1706$ \\
        3D & $30 \times 30 \times 3.2$ & $\left[ -0.52, 3.48 \right] \times \left[ -2, 2 \right] \times \left[ 0, 3.2 \right]$ & $0.008 \times 0.008 \times 0.08$ & $1.01$ & $1071 \times 1072 \times 40$ \\
        \hline
    \end{tabular}
\end{table*}

@@

\begin{figure}
    \centering
    \includegraphics[width=8cm]{figures/wz_avg_multi_contourf.png}
    \caption{Filled contour of the spanwise-averaged z-component of the vorticity ($-5 \leq w_z c / U_\infty \leq 5$) field after $80$ and $100$ time units of three-dimensional flow simulation with PetIBM for the snake cylinder with a cross-section at a $35$-degree angle of attack and Reynolds number $2000$.}
    \label{fig:wz_avg_3d}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=8cm]{figures/ux_uy_xz_plane.png}
    \caption{Filled contour of the streamwise velocity (top) and cross-flow velocity (bottom) behind the snake cylinder in the $x/z$ plane at $y/c=-0.2$ in the wake of the snake cylinder with a $35$-degree angle of attack at Reynolds number $2000$ after $100$ time-units of flow simulations. There are 52 contours from $-1.0$ to $1.0$. The grey area shows a projection of the snake cylinder in the $x/z$ plane. The solid black line defines the contour with $u_x = 0$ (top) and $u_y = 0$ (bottom).}
    \label{fig:ux_uy_xz_plane_3d}
\end{figure}
